---
title: "607-wk2-tb"
author: "Tyler Frankenberg"
date: "2/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(rvest)

```

## Importing csv files.

We start by importing the two .csv files containing our datapoints, as separate dataframes.


```{r read-csv-files}

population <- as.data.frame(read.csv("https://raw.githubusercontent.com/curdferguson/607-wk2-tb/cec949647cf86c56e60a073fe6d215d9c9daf1b0/population(2).csv"))

tb <- as.data.frame(read.csv("https://raw.githubusercontent.com/curdferguson/607-wk2-tb/cec949647cf86c56e60a073fe6d215d9c9daf1b0/wk2-tb-cases.csv"))

## rename the dataframe tb's columns
names(tb) <- c("country", "year", "cases")

```


## Joining the tables

We are not concerned with years for which we do not have both a population and a tb_cases datapoint, so we use an inner join to combine the two tables.

```{r join-tables}

tb_rates <- inner_join(tb, population)

```
## Calculate the tb rate

We calculate TB rate by mutating the "tb_rates" dataframe to include a column that divides cases by population for that country and year.

```{r calculate-rates}

tb_rates <- tb_rates %>%
  mutate(tb_rate = cases/population)

```


## Example analysis: two countries over time.

How did the tb rates in Afghanistan and Iraq change from year-to-year over the period from 1995 - 2015?  We can use a subset function to isolate the data for just the two countries in its own data frame, and then a simple ggplot to display the two countries' changing rates over time side-by-side.

```{r plot-data}

tb_rates_afiq <- subset(tb_rates, country == "Afghanistan" | country == "Iraq")

ggplot(tb_rates_afiq, aes(year, tb_rate, col=country)) +
  geom_line()

```
